**=====================================================**

# Machine Failure Analysis 
Noah MacGillivray | BIT-446 Assignment(T6) | Summer 2025 

Logistic Regression Classification 

**Environment:** Ubuntu, VS Code

**=====================================================**

## Introduction
This program imports 


## Required packages 
```{r}
# Load packages
library(readxl)
library(class)
library(caTools)
library(caret)
library(ggplot2)
library(lattice)
library(metrica)

cat("All packages loaded successfully!\n")
```

## Data Import and Initial Exploration 
The coffee shop data is read in, head/tail/summary are printed to
ensure proper import and view summary, and 
```{r}

# Read in sheet names
sheet_names <- excel_sheets("BIT-446-RS-T6-Lattes1.xlsx")
print(sheet_names)

all_sheets <- list()

for (sheet in sheet_names) {
    all_sheets[[sheet]] <- read_excel("BIT-446-RS-T6-Lattes1.xlsx", sheet = sheet) # nolint
}

print(all_sheets)

for (x in sheet_names) {
  cat("Sheet:", x, "\n")
  print(tail(all_sheets[[x]]))
  cat("\n")
}

```

## General impressions from summary 
